Stochastic Wind Power Forecasting

### 2020-01-28 LaTeX_code/paper/paper_1line.tex
more on Lamperti

Our model in (\ref{VtSDE}) is  a state-dependent diffusion formulation, that is the diffusion of the SDE clearly depends on the state.  We are interested in state independent diffusion as it is more tractable and numerically stable. To achieve that, we apply the Lamperti transform \autocite[98--100]{saso} to obtain the following state-independent diffusion SDE,
\begin{align}
    dZ_t&= \frac{- \theta_t (1+ \sin(Z_t) - 2p_t) + \alpha \theta_t \sin (Z_t)   }{\cos (Z_t)} \ dt + \sqrt{2 \alpha \theta_t} dW_t \\
    Z_0&=z_0 \,,
\end{align}
where the Lamperti trasform is given by $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $. We can see in Figure (\ref{fig:skew_after_lamperti}) the effects of the Lamperti transform on the forecast error data. The Lamperti transform has greatly reduced  the forecast error skewness and ensures that the process stays in the range $[0,1]$. Thus, in this case,  we can consider using a Gaussian density as a proxy transition density.
%---END SECTION 3---

>>>>

Our model (\ref{VtSDE}) for the forecast error has a diffusion term that depends on the state variable $V_t$. To estimate the unknown model parameters, a recommended technique is to modify the SDE (\ref{VtSDE}) by applying the so-called Lamperti transform (see \cite[40--41]{iacus1}, \cite{moma}, \cite[98--100]{saso}) to the process $V$ to obtain a SDE for the transformed process whose diffusion term does not depend anymore on the state variable. \\

We consider the following transformation 
\begin{equation}
Z_t = h(V_t) = \int_{\xi}^{x} \frac{1}{\sqrt{(u + p_t)(1 - u - p_t)}} \,du \Bigg{\vert}_{x = V_t} \,,
\end{equation}
where $\xi$ is an arbitrary point of the state space of the process $V$. The choice of $\xi = \frac{1}{2} - p_t$, where $p_t$ is a known input, leads to the process 
\begin{equation}
Z_t = h(V_t) = \arcsin (2(V_t + p_t) -1)\,,
\end{equation}
that, after applying It\^{o}'s formula on $h(V_t)$, gives the state-independent diffusion SDE
\begin{equation}
dZ_t = \left[  \frac{- \theta_t V_t}{\sqrt{(V_t + p_t)(1 - V_t - p_t)}} - \frac{1}{2} \frac{\left( \frac{1}{2} - (V_t + p_t)\right)}{(V_t + p_t)(1 - V_t - p_t)}  \right] \,dt + \sqrt{2 \theta_t \alpha}\, dW_t \,. \label{eq:stindepSDE}
\end{equation}
After replacing $V_t = \frac{\sin(Z_t) + 1}{2} - p_t$ in (\ref{eq:stindepSDE}), we obtain that the  process $Z$ satisfies the SDE
\begin{equation}
dZ_t = \frac{- \theta_t (\sin(Z_t) + 1 - 2 p_t) + \frac{\sin{Z_t}}{2}}{\cos{Z_t}} \,dt + \sqrt{2 \theta_t \alpha}\, dW_t \,. \label{eq:stindepSDE}
\end{equation}

%\begin{align}
%    dZ_t&= \frac{- \theta_t (1+ \sin(Z_t) - 2p_t) + \alpha \theta_t \sin (Z_t)   }{\cos (Z_t)} \ dt + \sqrt{2 \alpha \theta_t} dW_t \\
%    Z_0&=z_0 \,,
%\end{align}
%where the Lamperti trasform is given by $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.

We can see in Figure (\ref{fig:skew_after_lamperti}) the effects of the Lamperti transform on the forecast error data. The Lamperti transform has greatly reduced  the forecast error skewness and ensures that the process stays in the range $[0,1]$. Thus, in this case,  we can consider using a Gaussian density as a proxy transition density.
%---END SECTION 3---
